Article template&#58; A reproducible RMarkdown template for Elsevier's journal format with examples for beginners

### 2019-02-12 Paper/Manuscript.Rmd
Add data preprocessing

``` {r preprocessData, echo=FALSE, results="hide"}
# here is where all the preprocessing happens
#
# ensure you comment any sections or functions in sufficient detail that
# their purpose can be readily understood without reading through the detail 
# of the code itself
```

Keywords\:  
Key 1\; Key 2\; Key 3\; Key 4\; Key 5\; Key 6 (max)

>>>>

``` {r preprocessData, echo=FALSE, results="hide"}
## here is where all the preprocessing happens
# ensure you comment any sections or functions in sufficient detail that
# their purpose can be readily understood without reading through the detail 
# of the code itself

# rename columns and create a new variable
df <- df%>%
  rename(sepal_length = X1)%>%
  rename(sepal_width = X2) %>%
  rename(petal_length = X3) %>%
  rename(petal_width = X4)%>%
  rename(iris_class = X5) %>%
 mutate(length_ratio =  sepal_length/ petal_length)

# fit linear models for relationship between sepal and petal width for each iris type
df_lms = df %>% 
  group_by(iris_class) %>%
  do(fit_length = lm(sepal_length ~ petal_length, data = .))

# generate convenient filtering variables
df <- df%>%
  mutate(
    # identify if sepal length is more than 1 standard deviation above mean
    long_sepal = ifelse(sepal_length>= (mean(df$sepal_length) + sd(df$sepal_length)), "Yes", "No"),
    # identify if the petal width is a specified value, in this case exactly 1 or 2
    integer_petal = ifelse((petal_width >= 0.9 & petal_width<= 1.1) | (petal_width>= 1.9 & petal_width<=2.1),"Yes", "No")
  )

# characterize petal length into 5 categories in increments of 1.5 cm
df  = df%>%
  mutate(petel_length_level = ifelse(petal_length>= 6 & petal_length <7.5,"VeryHigh",0))%>%   
  mutate(petel_length_level = ifelse(petal_length>= 4.5 & petal_length <6,"High",petel_length_level))%>%  
  mutate(petel_length_level = ifelse(petal_length>= 3 & petal_length <4.5,"Medium",petel_length_level))%>%
  mutate(petel_length_level = ifelse(petal_length>= 1.5 & petal_length <3,"Low",petel_length_level))%>%
  mutate(petel_length_level = ifelse(petal_length <1.5,"VeryLow",petel_length_level))

# order new petal length categories correctly from Low to High
df = df%>%mutate(petel_length_level = as.factor(petel_length_level))
df$petel_length_level = factor(df$petel_length_level, levels=c("VeryLow","Low","Medium","High","VeryHigh"))
```

Keywords\:  
Keyword 1\; Keyword 2\; Keyword 3\; Keyword 4\; Keyword 5\; Keyword 6 (max)
